Convective heat transport in stratified atmospheres at low and high Mach number

### 2017-02-24 tex/nusselt_paper.tex
restructures experiment and results, need to finalize plots and put them in.

%%%%% Body of the paper
\section{Introduction}
\refstepcounter{section}
\label{sec:intro}
Convective dynamics transport energy in stellar and planetary atmospheres.
In these objects, flows are compressible and
feel the atmospheric stratification.  While in some systems this stratification is
negligible, it is significant in regions such as
the convective envelope of the Sun, which spans 14 density scale heights.
In the bulk of these systems, flows are at very low Mach Number (Ma), but numerical
constrainst have restricted most studies of compressible convection to high Ma.
These studies provide important insight to low temperature, high Ma regions
near the Sun's surface, but few fundamental
properties of low Ma convection which characterize deeper motions
are known.

@@

In RB convection, there exist two primary dynamical control parameters: 
the Rayleigh number (Ra, the ratio of
buoyant driving to diffusive damping) and the Prandtl number 
(Pr, the ratio of viscous to thermal
diffusivity). These numbers control two useful
measures of turbulence in the evolved solution:
the Reynolds
number (Re, the strength of advection to viscous diffusion)
and the Peclet number (Pe = Re Pr).  Stratified atmospheres
with negative entropy gradients are unstable to convection,
and the magnitude of that entropy gradient, or the superadiabaticity,
joins Ra and Pr as an important control parameter.  The
Ma of the evolved solution is primarily controlled by 
the \emph{superadiabatic excess}, $\epsilon$, which 
sets the scale of the atmospheric entropy gradient \cite{graham1975}.

@@

\section{Experiment} 
\refstepcounter{section}
\label{sec:experiment}
We examine the simplest stratified extension of RB by studying a
fluid composed of monatomic ideal gas particles with an adiabatic 
index of $\gamma = 5/3$ and whose equation of state is $P = R\rho T$. 
This is consistent with the approach used in earlier work.
The initial atmosphere is a plane-parallel polytrope in which 
the gravitational acceleration and conductive flux, 
$\bm{F}_{\text{cond,0}} = -\kappa \partial_z T_0$, do not vary with depth. To
achieve the latter condition, both $\kappa$ and $\partial_z T_0$ are constant.
Under these assumptions, satisfying hydrostatic 
equilibrium produces a stratification of
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(z_0 - z)^m, \\
T_0(z)    &= T_{t}(z_0 - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m = m_{\text{ad}} - \epsilon$ is the polytropic index.
The adiabatic polytropic index is $m_{\text{ad}} \equiv (\gamma-1)^{-1}$, and
the superadiabatic excess is $\epsilon$ which sets the scale 
of the entropy gradient ($\partial_z S_0 \propto -\epsilon$).
A significant advance of this work is the ability to study large 
and small values of $\epsilon$, as will be discussed.
Thermodynamic variables are nondimensionalized at the 
top of the atmosphere as  $P_0(L_z) = \rho_0(L_z) = T_0(L_z) = 1$, 
requiring $z_0 \equiv L_z + 1$ and $R = T_{t} = \rho_{t} = 1$.
By this choice, the non-dimensional length scale is the inverse 
temperature gradient scale and the timescale is the isothermal 
sound crossing time of this unit length.
The height $z$ increases upwards within $[0, L_{z}]$, 
where $L_{z} = e^{n_{\rho}/m} - 1$ is
determined by $n_\rho$ and $\epsilon$.
The characteristic timescale of convective dynamics
is related to the atmospheric buoyancy time, 
$t_{\text{b}} = \sqrt{L_z/g\epsilon}$, with $g = (m+1)$.
Throughout this letter, we use buoyancy time units and 
choose $n_{\rho} = 3$ such that the 
initial density varies by a factor of 20.
All atmospheres studied here have an aspect ratio of 4, 
such that $L_x = 4L_z$.

The atmospheric diffusivities are primarily controlled by
the non-dimensional Rayleigh number,
\begin{equation}
\text{Ra} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu\chi},
\end{equation}
where $\Delta S_0 = \epsilon\ln z_0$ is the entropy difference 
between the top and bottom boundaries, 
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat 
at constant pressure, $\nu$ is the kinematic viscosity 
(the viscous diffusivity), and $\chi$ is the thermal diffusivity.  
The relationship between the thermal and viscous diffusivities is
set by the Prandtl number, $\text{Pr} = \nu/\chi$.   
The dynamic viscosity, $\mu$, and the thermal conductivity,
$\kappa$, relate to their corresponding diffusivities such that 
$\nu \equiv \mu/\rho$ and $\chi \equiv \kappa/\rho$.  We take
$\mu$ and $\kappa$ to be constant with height.
As a result, $\text{Ra} \propto (\nu\chi)^{-1} \propto
\rho^2$.  The atmospheres studied here with $n_{\rho} = 3$ 
experience an increase in Ra 
by a factor of 400 across the domain.  This formulation leaves Pr
constant throughout the depth of the atmosphere. 
In this letter we impose $\text{Pr} = 1$ and specify Ra at 
the top of the domain ($z = L_z$).

While holding $n_\rho$ and Pr constant, 
the primary control parameters of convection are $\epsilon$
and Ra.  We decompose our atmosphere into the background 
polytrope ($\ln\rho_{0}, T_{0}$) and the fluctuations
about that background ($\bm{u}, \ln\rho_{1}, T_{1}$),
which can be large.
The scaling of the entropy gradient with $\epsilon$
is reflected in the evolved values of these fluctuations.  For small $\epsilon$,
evolved values scale as
$T_1/T_0 \propto \rho_{1}/\rho_{0} \propto$ Ma$^{2} \propto \epsilon$,
as shown in Fig. \ref{fig:ma_v_eps}.

@@

We evolve the Fully Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \grad\cdot\bm{u} 
    = -\bm{u}\cdot\grad\ln\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - 
&\nu\grad\cdot\lilstressT - \lilstressT\cdot\grad\nu = \\
&-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + 
\nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.
    \grad^2 T + \grad T\cdot\grad\chi\right) = \\
	&-\bm{u}\cdot\grad T - (\gamma-1)T\grad\cdot{\bm{u}} \\
	&+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +
	\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
with the viscous stress tensor given by
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + 
\frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\grad\cdot\bm{u}\right).
	\label{eqn:stress_tensor}
\end{equation}
Taking an inner product of
(\ref{eqn:momentum_eqn}) with $\bm{u}$ and adding it to 
(\ref{eqn:energy_eqn}) reveals the full energy equation,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
	\label{eqn:energy_eqn_full}
\end{equation}
where
$
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}}
$
is the convective flux and $\bm{F}_{\text{cond}} = -\kappa \grad T$
is the conductive flux.
The individual contributions to $\bm{F}_{\text{conv}}$ are the enthalpy flux, 
$\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$;
the kinetic energy flux, 
$\bm{F}_{\text{KE}} \equiv \rho|\bm{u}|^2\bm{u}/2$;
the potential energy flux,
$\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ (with $\phi \equiv -gz$);
and the viscous flux, 
$\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$, and each 
must be considered. 
Understanding how these fluxes interact  
is crucial in characterizing convective heat transport.

The atmosphere is contained between two 
impenetrable, stress free, fixed temperature boundaries at
the top and bottom of the domain such that 
$w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. The domain
is horizontally periodic. We utilize the 
Dedalus\footnote{http://dedalus-project.org/} \cite{burns&all2016} pseudospectral framework 
 to time-evolve  
(\ref{eqn:continuity_eqn})-(\ref{eqn:energy_eqn}) 
using an implicit-explicit, third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
Physical grid sizes range from
96x384 grid points at the lowest values of 
Ra to 1152x4608 grid points at Ra $\geq 10^{7}$. 
By using IMEX timestepping, we implicitly step the 
stiff linear acoustic wave contribution and are able to
efficiently study flows at moderate ($\approx 1$) and very low ($\approx 10^{-4}$)
Ma (Fig. \ref{fig:ma_v_eps}b).  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include variable
$\nu$ and $\chi$, and we follow the approach there; 
this IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible 
Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

The efficiency of convection is quantified by the Nusselt number.  
Nu is well-defined in RB convection
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, and we use
a modified version of a traditional stratified Nusselt number \cite{graham1975,hurlburt&all1984},
\begin{equation}
\text{Nu} \equiv \frac{\langle F_{\text{conv, z}} + F_{\text{cond, z}} - F_{\text{A}}\rangle}
{\langle F_{\text{cond, z}} - F_{\text{A}}\rangle} 
= 1 + \frac{\langle F_{\text{conv, z}}\rangle}{\langle F_{\text{cond, z}} - F_{\text{A}} \rangle}
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv, z}}$ and $F_{\text{cond, z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
respectively and $\langle \rangle$ implies a volume average.  
$F_{\text{A}} \equiv -\langle\kappa\rangle \partial_z T_{\text{ad}}$ 
is the adiabatic conductive flux and
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ 
for an ideal gas in hydrostatic equilibrium.  Here we specify
$\langle \kappa \rangle = \langle (\rho_{1}/\rho_{0})\kappa_0 \rangle$.
At large values of $\epsilon$, $\kappa$ evolves significantly as the
density profile evolves.  At low $\epsilon$, $\langle\kappa\rangle \approx \kappa_0$.
In the Boussinesq approximation, under which $\grad S = 0$ only when 
$\grad T = 0$, this definition reduces the the traditional definition
of the Nusselt number \cite{otero&all2002, johnston&doering2009}.

@@

\begin{figure}[b]
\includegraphics[width=3.4375in]{./figs/snapshots_fig.png}
\caption{Characteristic entropy fluctuations in evolved flows. 
The time- and horizontally-averaged profile is removed in all cases.  (a) At high
$\epsilon$, shock systems form near the upper downflow lanes ($x \approx 45, z \approx 15-19$)
at sufficiently high Ra.
Shock-heated fluid then flows into the downflows as the shocks propagate across upflows.
(b) At low $\epsilon$ but at the same Ra, shock systems are absent, 
but otherwise the dynamics are similar.  (c) As Ra is increased, downflows no longer span
the entirety of the domain and individual small eddies are responsible for carrying the flux.
\label{fig:entropy_snapshots} }
\end{figure}

We evolve initial value problems in which $T_1$ is
filled with infinitesimal, random white noise compared to $T_0$ and $\epsilon$.
We filter the noise spectrum in coefficient space, such that 25\% of the coefficients
have power.
Solutions were time-evolved until a long average of Nu showed little
variance with depth. By performing a linear stability analysis, 
we determined that the onset of convection
occurs at $\text{Ra}_c = \{10.06, 10.97, 10.97\}$ for $\epsilon = \{0.5, 10^{-4}, 10^{-7}\}$ respectively.  
We studied Rayleigh
numbers from values at onset up to nearly $10^6$Ra$_c$ for $\epsilon = \{0.5, 10^{-4}\}$ 
and up to $10^3$Ra$_c$ for $\epsilon = 10^{-7}$.

At large Ma ($\epsilon = 0.5$), shock systems form in the upper atmosphere near downflow lanes 
(Fig. \ref{fig:entropy_snapshots}a) once Ra is sufficiently large.
These shocks propagate through upflow regions.  
Such systems were reported in
both two \cite{cattaneo&all1990} and three \cite{malagoli&all1990} dimensional polytropic simulations previously.
These shocks heat material entering the downflows, affecting the dynamics and heat transport
of these systems.

Low Ma flows ($\epsilon = 10^{-4}$)
have similar bulk thermodynamic structures (Fig. \ref{fig:entropy_snapshots}b)
to high Ma flows.
As Ra is increased to large values 
(Fig. \ref{fig:entropy_snapshots}c), thermodynamic structures 
no longer span the whole domain but rather break up into 
small eddies which traverse the domain multiple
times before diffusing.  
While it has been suggested that pressure forces 
cause symmetry breaking in up- and downflows
\cite{hurlburt&all1984}, at low $\epsilon$ this 
effect seems to be secondary to flows obeying mass conservation as they traverse
the stratified medium.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/re_and_nu_v_Ra.png}
\caption{Time-averaged vertical flux profiles ($F_{\text{z}}$)
for (a) high and (b) low Ma flows at $\text{Ra }= 10^6$.  
All fluxes are defined as in (\ref{eqn:energy_eqn_full}) and 
normalized by $F_{\text{ref}} - F_{\text{A}}$, as in (\ref{eqn:nusselt}).
The dashed lines correspond to the
enthalpy flux (orange, positive) and kinetic energy flux (purple, negative).  
The grey dash-dot line is the
viscous flux and the green dotted line is the conductive flux with the 
adiabatic contribution removed. The potential energy flux is negligible and is not shown.
In (a), the viscous flux is negligible and is not shown.  The solid black line is
Nu, the properly normalized sum of all the fluxes \label{fig:re_and_nu_v_Ra} }
\end{figure}

At large enough values of Ra for shocks to form,
high Ma flows exhibit two 
local maxima in the enthalpy flux and kinetic energy flux (Fig. \ref{fig:flux_profiles}a).
Shock-heated fluid parcels sometimes gain vorticity as
they sink into the lower atmosphere.
This creates deep, rapidly-rotating regions
of mixing which persist for many overturn times. These ``spinners'' appear to
influence the dynamics, but their contributions are unclear.

At low Ma, only the deep maximum in enthalpy and kinetic energy fluxes
is present (Fig. \ref{fig:flux_profiles}b).  
Our choice of fixed-temperature boundary conditions allows the flux at the 
boundaries to vary, so many runs at $\text{Ra }> 10^5$
and $\epsilon = 10^{-4}$ exhibit states in which the flux 
entering the system at the bottom of the atmosphere 
exceeds that which leaves at the top.  
These systems are punctuated by states of vigorous shearing, similar to those previously
reported in two-dimensional RB convection \cite{goluskin&all2014}.  
%In these shearing states, flux entering the lower atmosphere approaches $F_\text{A}$,
%allowing excess energy to escape through the upper boundary.  
During shearing states, convective transport is suppressed and Nu diminishes while
excess energy exits the system through the upper boundary.
A proper long-term average over shearing
and non-shearing states retrieves an invariant Nu profile throughout the depth
of the atmosphere. 
These shearing states will be covered in more detail in a future paper.

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/density_v_ra.png}
\caption{Variation of Nu as Ra increases at high and low $\epsilon$. 
At high $\epsilon$ (purple circles), 
a clear transition from the subsonic to supersonic regime is evident in the scaling
of Nu with Ra at $\text{Ra} \approx 10^4$.  In the low $\epsilon$ regime (green diamonds and yellow squares), 
our observed Nu scalings collapse onto a similar line which is
indistinguishable from a $\text{Ra}^{2/7}$ scaling, 
as observed in RB convection \cite{johnston&doering2009}.  Error bars are determined
 by the square root of the variance of the time-averaged Nu profile with depth and 
 indicate whether or not a solution is well-converged.
\label{fig:nrho_v_ra} }
\end{figure}

>>>>

%%%%% Body of the paper
\section{Introduction}
\refstepcounter{section}
\label{sec:intro}
Convective dynamics transport energy in stellar and planetary atmospheres.
In these objects, flows are compressible and
feel the atmospheric stratification.  While in some systems this stratification is
negligible, it is significant in regions such as
the convective envelope of the Sun, which spans 14 density scale heights.
In the bulk of these systems, flows are at very low Mach Number (Ma), but numerical
constrainst have restricted most studies of compressible convection to high Ma.
These studies have provided important insight into the low temperature, high Ma region
near the Sun's surface, but few fundamental
properties of low Ma convection which characterize deeper motions
are known.

@@

In RB convection, there exist two primary dynamical control parameters: 
the Rayleigh number (Ra, the ratio of
buoyant driving to diffusive damping) and the Prandtl number 
(Pr, the ratio of viscous to thermal
diffusivity). These numbers control two useful
measures of turbulence in the evolved solution:
the Reynolds
number (Re, the strength of advection to viscous diffusion)
and the Peclet number (Pe = Re Pr).  Stratified atmospheres
with negative entropy gradients are unstable to convection.
The magnitude of the unstable entropy gradient, or the superadiabaticity,
joins Ra and Pr as an important control parameter.  This 
\emph{superadiabatic excess} \cite{graham1975}, $\epsilon$, which 
sets the scale of the atmospheric entropy gradient
primarily controls the Ma of the evolved solution.

@@

\section{Experiment} 
\refstepcounter{section}
\label{sec:experiment}
We examine a monatomic ideal gas with an adiabatic index of
$\gamma = 5/3$ whose equation of state is $P = R\rho T$.
This is consistent with the approach used in earlier work
and is the simplest stratified extension of RB.
We study atmospheres which are initially polytropically stratified,
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(1 + L_z - z)^m, \\
T_0(z)    &= T_{t}(1 + L_z - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m$ is the polytropic index and $L_z$ is the depth of the atmosphere.
The height coordinate, $z$, increases upwards in the range $[0, L_z]$.  We
specify the depth of the atmosphere, $L_z = e^{n_{\rho}/m} - 1$, by choosing
the number of density scale heights, $n_{\rho}$, it spans.
Throughout this letter we set $n_{\rho} = 3$.  Satisfying hydrostatic
equilibrium sets the value of gravity, $g = R T_t (m + 1)$, which is
constant with depth.  We study two-dimensional atmospheres with aspect
ratios of 4 where the $x$ coordinate has the range $[0, 4L_z]$.

%
%We examine the simplest stratified extension of RB by studying a
%fluid composed of monatomic ideal gas particles with an adiabatic 
%index of $\gamma = 5/3$ and whose equation of state is $P = R\rho T$. 
%This is consistent with the approach used in earlier work.
%The initial atmosphere is a plane-parallel polytrope in which 
%the gravitational acceleration and conductive flux, 
%$\bm{F}_{\text{cond,0}} = -\kappa \partial_z T_0$, do not vary with depth. To
%achieve the latter condition, both $\kappa$ and $\partial_z T_0$ are constant.
%Under these assumptions, satisfying hydrostatic 
%equilibrium produces a stratification of
%\begin{equation}
%\begin{split}
%\rho_0(z) &= \rho_{t}(z_0 - z)^m, \\
%T_0(z)    &= T_{t}(z_0 - z),
%\label{eqn:polytrope}
%\end{split}
%\end{equation}
%where $m = m_{\text{ad}} - \epsilon$ is the polytropic index.
%The adiabatic polytropic index is $m_{\text{ad}} \equiv (\gamma-1)^{-1}$, and
%the superadiabatic excess is $\epsilon$ which sets the scale 
%of the entropy gradient ($\partial_z S_0 \propto -\epsilon$).
%A significant advance of this work is the ability to study large 
%and small values of $\epsilon$, as will be discussed.
%Thermodynamic variables are nondimensionalized at the 
%top of the atmosphere as  $P_0(L_z) = \rho_0(L_z) = T_0(L_z) = 1$, 
%requiring $z_0 \equiv L_z + 1$ and $R = T_{t} = \rho_{t} = 1$.
%By this choice, the non-dimensional length scale is the inverse 
%temperature gradient scale and the timescale is the isothermal 
%sound crossing time of this unit length.
%The height $z$ increases upwards within $[0, L_{z}]$, 
%where $L_{z} = e^{n_{\rho}/m} - 1$ is
%determined by $n_\rho$ and $\epsilon$.
%The characteristic timescale of convective dynamics
%is related to the atmospheric buoyancy time, 
%$t_{\text{b}} = \sqrt{L_z/g\epsilon}$, with $g = (m+1)$.
%Throughout this letter, we use buoyancy time units and 
%choose $n_{\rho} = 3$ such that the 
%initial density varies by a factor of 20.
%All atmospheres studied here have an aspect ratio of 4, 
%such that $L_x = 4L_z$.
%

Convective dynamics are controlled by the superadiabaticity
of the atmospheres as well as the atmospheric diffusivities.
The superadiabaticity, or the magnitude of the (negative) 
entropy gradient, is set by 
the superadiabatic excess, $\epsilon = m_{ad} - m$, where 
$m_{ad} = (\gamma - 1)^{-1}$.
The atmospheric thermal diffusivity, $\chi$,
and kinematic viscosity (or viscous diffusivity), $\nu$,
are determined by the Rayleigh number (Ra) and the Prandtl
number (Pr).  We set $\text{Pr } = \nu/\chi = 1$ throughout
this work. The polytropic initial conditions are in
thermal equilibrium, $\kappa_0 \partial_z T_0 =$ const,
so $\kappa_0 = \chi \rho_0$ must be constant. To keep
Pr constant with height, we set $\chi = \chi_t/\rho_0$
and $\nu = \nu_t/\rho_0$.  Under these constraints, the
diffusivity profiles are specified by choosing Ra
at the top of the domain,
\begin{equation}
\text{Ra}_{\text{top}} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu_t\chi_t},
\end{equation}
where $\Delta S_0 = \epsilon\ln z_0$ is the entropy difference 
between the top and bottom boundaries and
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat 
at constant pressure. The profiles of $\nu$ and $\chi$ are constant with
time, and Ra at the bottom of the atmosphere is greater than Ra$_{\text{top}}$
by a factor of $e^{2n_{\rho}}$.  This formulation leaves
the thermal conductivity, $\kappa = \rho\chi$ and
the dynamic viscosity, $\mu = \rho\chi$, free to evolve
as the density profile changes.

We decompose our thermodynamic variables such that $T = T_0 + T_1$ and
$\ln\rho = (\ln\rho)_0 + (\ln\rho)_1$, and the velocity $\bm{u}$ has no
background component.    We evolve the Fully Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \grad\cdot\bm{u} 
    = -\bm{u}\cdot\grad\ln\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - 
&\nu\grad\cdot\lilstressT - \lilstressT\cdot\grad\nu = \\
&-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + 
\nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.
    \grad^2 T + \grad T\cdot\grad\chi\right) = \\
	&-\bm{u}\cdot\grad T - (\gamma-1)T\grad\cdot{\bm{u}} \\
	&+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +
	\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
with the viscous stress tensor given by
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + 
\frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\grad\cdot\bm{u}\right).
	\label{eqn:stress_tensor}
\end{equation}
Taking an inner product of
(\ref{eqn:momentum_eqn}) with $\bm{u}$ and adding it to 
(\ref{eqn:energy_eqn}) reveals the full energy equation,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
	\label{eqn:energy_eqn_full}
\end{equation}
where
$
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}}
$
is the convective flux and $\bm{F}_{\text{cond}} = -\kappa \grad T$
is the conductive flux.
The individual contributions to $\bm{F}_{\text{conv}}$ are the enthalpy flux, 
$\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$;
the kinetic energy flux, 
$\bm{F}_{\text{KE}} \equiv \rho|\bm{u}|^2\bm{u}/2$;
the potential energy flux,
$\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ (with $\phi \equiv -gz$);
and the viscous flux, 
$\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$, and each 
must be considered. 
Understanding how these fluxes interact  
is crucial in characterizing convective heat transport.

We utilize the 
Dedalus\footnote{http://dedalus-project.org/} \cite{burns&all2016} pseudospectral framework 
 to time-evolve  
(\ref{eqn:continuity_eqn})-(\ref{eqn:energy_eqn}) 
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
Physical grid sizes range from
96x384 grid points at the lowest values of 
Ra to 1536x6144 grid points at Ra $> 10^{7}$. 
By using IMEX timestepping, we implicitly step the 
stiff linear acoustic wave contribution and are able to
efficiently study flows at moderate ($\approx 1$) and very low ($\approx 10^{-4}$)
Ma.  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include
$\nu$ and $\chi$ which vary with depth, and we follow the approach there.
This IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible 
Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

We impose
impenetrable, stress free, fixed temperature boundary conditions at
the top and bottom of the domain such that 
$w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. 
$T_1$ is initially filled with
random white noise whose magnitude is infinitesimal
compared to $T_0$ and $\epsilon$.
We filter this noise spectrum in coefficient space, such that 25\% of the coefficients
have power. We non-dimensionalize our computational domains by setting
all thermodynamic variables to unity at $z = L_z$ and
$R = T_t = \rho_t = 1$.  By this choice, the non-dimensional
grid length scale is the inverse temperature gradient scale and the 
simulation timescale is the isothermal sound crossing time, $\tau_I$ of this unit length.
Meaningful convective dynamics occur on timescales of the atmospheric buoyancy time,
$t_b = \tau_I \sqrt{L_z/g\epsilon}$, and we will report our results in terms of
buoyancy times.

The efficiency of convection is quantified by the Nusselt number (Nu).  
Nu is well-defined in RB convection
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, and we use
a modified version of a traditional stratified Nusselt number \cite{graham1975,hurlburt&all1984},
\begin{equation}
\text{Nu} \equiv \frac{\langle F_{\text{conv,z}} + F_{\text{cond,z}} - F_{\text{A}}\rangle}
{\langle F_{\text{cond,z}} - F_{\text{A}}\rangle} 
= 1 + \frac{\langle F_{\text{conv,z}}\rangle}{\langle F_{\text{cond,z}} - F_{\text{A}} \rangle}
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv, z}}$ and $F_{\text{cond, z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
respectively and $\langle \rangle$ implies a volume average.  
$F_{\text{A}} \equiv -\langle\kappa\rangle \partial_z T_{\text{ad}}$ 
is the adiabatic conductive flux and
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ 
for an ideal gas in hydrostatic equilibrium.  Here we specify
$\langle \kappa \rangle = \langle \rho\chi \rangle$, which is nearly
$\kappa_0$ when $\epsilon$ is small but can change appreciably for large
values of $\epsilon$.
In the Boussinesq approximation, under which $\grad S = 0$ only when 
$\grad T = 0$, this definition reduces the the traditional definition
of the Nusselt number \cite{otero&all2002, johnston&doering2009}.

@@

Solutions were time-evolved until a long time average of the fluxes
showed little
variance with depth. By performing a linear stability analysis, 
we determined that the onset of convection
occurs at $\text{Ra}_c = \{11.15, 10.06, 10.97, 10.97\}$ for $\epsilon = \{1.0, 0.5, 10^{-4}, 10^{-7}\}$ respectively.  
We studied Rayleigh
numbers from values at onset to values $\geq 10^6$Ra$_c$ for $\epsilon = 0.5$, up to
roughly $10^5$Ra$_c$ at $\epsilon = 10^{-4}$, and 
up to $10^3$Ra$_c$ for $\epsilon = \{1, 10^{-7}\}$.

We find that the Mach number is a strong function of $\epsilon$ and a weak function of 
Ra, as can be seen in Fig. \ref{fig:ma_v_eps}.  We find that Ma $\propto \epsilon^{1/2}$,
which we expect at low $\epsilon$ but which seems to persist even as $\epsilon \rightarrow m_{ad}$.
Furthermore, the evolved thermodynamic variables, $T_1/T_0, \rho_1/\rho_0 \propto \epsilon$,
are very small for low Ma flows, but are order 1 when $\epsilon$ and Ma approach 1.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/snapshots_fig.png}
\caption{Characteristic entropy fluctuations in evolved flows. 
The time- and horizontally-averaged profile is removed in all cases.  (a) At high
$\epsilon$, shock systems form near the upper downflow lanes ($x \approx 45, z \approx 15-19$)
at sufficiently high Ra.
Shock-heated fluid then flows into the downflows as the shocks propagate across upflows.
(b) At low $\epsilon$ but at the same Ra, shock systems are absent, 
but otherwise the dynamics are similar.  (c) As Ra is increased, downflows no longer span
the entirety of the domain and individual small eddies are responsible for carrying the flux.
\label{fig:entropy_snapshots} }
\end{figure}

At large Ma ($\epsilon = 0.5$), shock systems form in the upper atmosphere near downflow lanes 
(Fig. \ref{fig:entropy_snapshots}b,c) once Ra is sufficiently large.
These shocks propagate through upflow regions.  
Such systems were reported in
both two \cite{cattaneo&all1990} and three \cite{malagoli&all1990} dimensional polytropic simulations previously.
These shocks heat material entering the downflows, affecting the dynamics and heat transport
of these systems.  As Ra is increased to large values 
(Fig. \ref{fig:entropy_snapshots}c), thermodynamic structures 
no longer span the whole domain but rather break up into 
small eddies which traverse the domain multiple
times before diffusing.

Low Ma flows ($\epsilon = 10^{-4}$)
have similar bulk thermodynamic structures (Fig. \ref{fig:entropy_snapshots}a)
to high Ma flows.
While it has been suggested that pressure forces 
cause symmetry breaking in up- and downflows
\cite{hurlburt&all1984}, at low $\epsilon$ this 
effect seems to be secondary to flows obeying mass conservation as they traverse
the stratified medium.  
Our choice of fixed-temperature boundary conditions allows the flux at the 
boundaries to vary, so many runs at $\text{Ra }> 10^4$
and $\epsilon = 10^{-4}$ exhibit long-term states of
flux disequilibrium.  Roll solutions such as those pictured in Fig. \ref{fig:entropy_snapshots}a
are punctuated are punctuated by states of vigorous shearing, similar to those previously
reported in two-dimensional RB convection \cite{goluskin&all2014}.  
During the roll states, the flux at the lower boundary exceeds that at the upper boundary
and energy is stored in the system.
During shearing states, convective transport is suppressed and Nu diminishes while
the flux at the upper boundary is higher than the flux at the lower boundary.  Long term
averages over these shearing and non-shearing states produce a flat flux profile which
can be sensibly analyzed.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/re_and_nu_v_Ra.png}
\caption{
Variation of Nu as Ra increases at high and low $\epsilon$. 
At high $\epsilon$ (purple circles), 
a clear transition from the subsonic to supersonic regime is evident in the scaling
of Nu with Ra at $\text{Ra} \approx 10^4$.  In the low $\epsilon$ regime (green diamonds and yellow squares), 
our observed Nu scalings collapse onto a similar line which is
indistinguishable from a $\text{Ra}^{2/7}$ scaling, 
as observed in RB convection \cite{johnston&doering2009}.  Error bars are determined
 by the square root of the variance of the time-averaged Nu profile with depth and 
 indicate whether or not a solution is well-converged.
}
\end{figure}

@@

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/density_v_ra.png}
\caption{\label{fig:nrho_v_ra} Here's how density stratification changes.}
\end{figure}
